Shading between two lines - ggplot
========================================================

First one to say geom_ribbon loses. I was plotting some data for a colleague, had two lines (repeated experiment) per person (time on the x axis) facetted by id, I thought it'd be nice to shade the area between the two lines so that when they were deviating you'd see a large shaded area, and when they were close there would be little shading, just to aid the visual of the separation between repeats. I thought this would be trival, and geom_ribbon would do the trick, alas, some of the lines crossed so that didn't pan out. Ignoring the 'experiment/order' variable and reordering the data to track the max and min values at each time point doesn't work, because when they cross you end up with a box around the cross, rather the inside region being shaded.

<!-- more -->

I did think this would have been done before, but I couldn't find anything that resuable. There was [this blog post](http://learnr.wordpress.com/2009/10/22/ggplot2-two-color-xy-area-combo-chart/) but like user Aniko mentioned in the comments, that was essentially finding the midpoints by hand, which seemed a bit clumsy (not saying the below is any better). Aniko's solution used a package [gpclib](http://cran.r-project.org/web/packages/gpclib/index.html) to create polygons for each block of colour, which was about where I got to when looking for a ggplot strategy. I played around a bit and couldn't get the gpc.poly data to work with ggplot, so put together a couple of small functions to essentially do the same thing. 

So here's the code and output. 

```{r}
# load data
library(ggplot2)
library(RCurl)
library(gridExtra)
library(plyr)
theme_set(theme_bw())

dat <- read.csv(textConnection(getURL("https://raw.githubusercontent.com/nzcoops/datasets/master/shading_two_lines")))
h(dat)

# this block is run within each person 
# essentially it creates a duplicate of all rows bar the first and last two
# and adds a grouping variable to the end
# that way every 4 rows are will be the coords for a polygon

mperson <- function(x){
  x <- x[order(x$time),]
  y <- x[-c(1,2,nrow(x)-1,nrow(x)), ]
  x <- rbind(x,y)
  x <- x[order(x$time),]
  x$group <- rep(letters[1:(nrow(x)/4)], each=4)
  return(x)
}
dat2 <- ddply(dat, .(id), mperson)
h(dat)

# this block is run within each person and 'block (group)' of 4 rows (each polygon)
# essentially this is to get the rows in the correct order, so that the geom_polygon function
# can work clockwise to construct the polygons the correct way
mgroup <- function(x){
  x <- x[order(x$bgl), ]
  left <- x[x$time == min(x$time), ]
  right <- x[x$time == max(x$time), ]
  if(all(left$order == right$order)){
    left <- left[order(left$bgl, decreasing = T), ]
    right <- right[order(right$bgl, decreasing = F), ]
    return(rbind(left, right))
  } else {
    return(x[order(x$time), ])
  } 
}
dat2 <- ddply(dat2, .(id,group), mgroup)
h(dat)
```
## And here's the plot
```{r shading_plot1, fig.width=10, fig.height=5}
ggplot(dat, aes(x=time, y=bgl, group=order)) + 
  geom_line(aes(colour=factor(order))) +
  geom_point(aes(colour=factor(order))) +
  geom_polygon(data=dat2, aes(y=bgl, group=group), alpha=0.3) +
  facet_wrap( ~ id)
```

I wrote this post in [RStudio](http://rstudio.org/download/) using the [R Markdown](http://www.rstudio.org/docs/authoring/using_markdown) language and then [knitr](http://yihui.name/knitr/) to turn in into markdown (.md), and then [pandoc](http://johnmacfarlane.net/pandoc/) to turn it into html. The original file is available here on [github](https://raw.githubusercontent.com/nzcoops/blog_code/master/shading_between_the_lines.Rmd).

system("pandoc -s shading_between_the_lines.md -o shading_between_the_lines.html")

As an aside, the mgroup function might seem like overkill, but it was a bit tricky, as when the lines cross you have to be careful to get the right 'hourglass' orientation, either vertical or horizontal.

```{r}
dat <- data.frame(x = c(10,10,20,20),
                  y = c(3,4,5,2),
                  order = c(1,2,1,2)
)

a <- ggplot(dat, aes(x=x, y=y)) + 
  geom_line(aes(group=order)) +
  geom_point(aes(group=order)) +
  geom_polygon(aes(x=x, y=y), fill="red", alpha=0.2) 

dat <- data.frame(x = c(10,10,20,20),
                  y = c(3,4,2,5),
                  order = c(1,2,1,2)
)

b <- ggplot(dat, aes(x=x, y=y)) + 
  geom_line(aes(group=order)) +
  geom_point(aes(group=order)) +
  geom_polygon(aes(x=x, y=y), fill="red", alpha=0.2) 

dat <- data.frame(x = c(10,20,10,20),
                  y = c(3,4,5,2),
                  order = c(1,2,2,1)
)

c <- ggplot(dat, aes(x=x, y=y)) + 
  geom_line(aes(group=order)) +
  geom_point(aes(group=order)) +
  geom_polygon(aes(x=x, y=y), fill="red", alpha=0.2) 
```

```{r shading_plot2, fig.width=15, fig.height=5}
grid.arrange(a,b,c, nrow=1)
```
